Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2017-04-20 2017-Xudong-SMT/submit-ICTAC2017.tex
Reduce the pages of the paper.
Main content simplified
1.remove the detail description of modification on the sketch in the head of the Section 5.
2.simplify the subsection "Management of assignments" remove the initialization of state variables.
3.in subsection "Interaction with Z3" simplify the detail, remove the time point declaring and translating the presentation.
4.in "Translating assignments into predicate terms" shorten the description, add an example about the translation.

\begin{figure}[h]

  \includegraphics[width=6cm]{XFIG/Enable1.eps}
%  \hspace{0.5cm}
  \includegraphics[width=6cm]{XFIG/Enable2.eps}
  \caption{Two pNet encodings for the Lotos Enable operator P>>Q }  \label{schema:enable-pnets}

\end{figure}

@@

\TODO{To be re-written}

In fact we give two encodings of this operator: the first one is
``state oriented'', in the sense that the controller has two states,
indicating where the control point is in the P>>Q process expression,
either in P or in Q. The transition between these occurs when process
P performs a $\delta(x)$ action, transferring the control to Q, and
sending a value $x$ at the same time. Process Q must be ready to
``accept'' this value through an $acc(x)$ action. The controller C1,
and the 3 synchronization vectors of the Enable1 pNet should be easy
to understand. Note 
that the first vector, transmitting an action $a1$ of P, can only
occur if $a1$ is not a $\delta$.

The second encoding is ``data oriented'', meaning that the state of
the node controller is encoded in the value of the pLTS state
variable(s). Here we have a single state variable $s_0$, initialized
as $0$. Its value is ``0'' when the control is on P, and ``1'' when the
control is on Q. The transitions in the pLTS have the more general
syntax: ``<action-expr> [guard] \{sequence of assignments\}''. Guards
and assignments can only use the values of variables of their source
state and values from input variables in the action expression; and a
transition can only assign variables of its target state. 
The second vector of the node ``Enable2'' also shows a
$\underline{\delta(y_2)}$ result action, where the underlined action
stands for an internal (already synchronized) action, that cannot be
further synchronized at upper levels. This is a straightforward
 generalization of the notion of internal actions, that will be
convenient for observing internal events during model-checking.

@@

The preceding definition relies on the auxiliary functions below:

\TODO{we do not use the Sort in the implementation of synch vector
  matching. May be we could remove it from here, but it will not gain
  much...}

\begin{definition}[Sorts, Holes, Leaves of pNets]
  \begin{itemize}
  \item The sort of a pNet is its signature, i.e. the set of actions it can
perform. In the definition of sorts, we do not need to distinguish
input variables (that specify the dataflow within LTSs), so for
computing LTS sorts, we use a substitution operator\footnote{$\subst{y_k\gets x_k}^{k\in K}$ is the parallel substitution 
operation.} to remove the
\emph{input marker} of variables. Formally:
\[
\begin{array}{l}
\Sort(\mylangle S,s_0, \to\myrangle) = \{\alpha\subst{x \gets ?x| 
x\in\symb{iv}(\alpha)}|s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
		J}\rangle} s'\in \to \} \\ \Sort(\mylangle
\set{\pNet}, \set{\symb{S}}, \set{\symb{SV}}\myrangle) =\{\alpha'_k |\, \alpha_j^{j\in J_k}\to\alpha'_k\in\set{\symb{SV}}\}
\end{array}
\]

\item
The set of holes of a pNet is defined inductively; the sets of holes
in a pNet node and its subnets are all disjoint:
  \[\begin{array}{l}
\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset \\
\Holes(\mylangle \pNet_i^{i\in I}\!,S_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) =J\cup{\displaystyle \bigcup_{i\in I}\Holes(\pNet_i)}\\
\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  \Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
\end{array}\]
\item
The set of leaves of a pNet is the set of all pLTSs occurring in the structure, defined inductively as:
\[\begin{array}{l}
\Leaves(\mylangle S,s_0, \to\myrangle) \!=\! \{ \mylangle S,s_0, \to\myrangle \}\\
\Leaves(\mylangle \pNet_i^{i\in I}\!,S_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) = {\displaystyle \bigcup_{i\in I}\Leaves(\pNet_i)}
\end{array}\]
\end{itemize}
\end{definition}

A pNet $Q$ is \emph{closed} if it has no hole: $\Holes(Q)=\emptyset$; else it
is said to be \emph{open}.

@@

\begin{example}\emph{An open-transition.}
  \label{OT:ccs-composed}
\TODO{: change for an OT of new Lotos example, may be only one or two
  of them ?}
The \texttt{CCS} pNet of Fig. \ref{schema:ccs-pnets} has 2 controllers and 2 holes. We show its full open Automaton in section \ref{section:full-result}, it has 16 open transitions. We detail 3 of them here:

 \smallskip
 $  OT_1  = \openrule{
                       0 \xrightarrow{a}_{C_1} 1 ~~
                       0 \xrightarrow{b}_{C_2} 1  ~~
                       \{\}               ~~
                 [ a\neq l \land b\neq r \land a=?ch(v) \land b=!ch(v) ]              
                      }
     {\ostate{00} \xrightarrow{\underline{\tau}} \ostate{11}}$

 $ OT_3  = \openrule{
                       0 \xrightarrow{a}_{C_1} 1 ~~
                       \{\}               ~~
                 [ a\neq l  ]              
                      }
     {\ostate{10} \xrightarrow{a} \ostate{10}} $

$  OT_{11}  = \openrule{
                          1 \xrightarrow{l}_{C_1} 1 ~~
                       0 \xrightarrow{b}_{C_2} 1 ~~
                       \{ \xrightarrow{hb_P}_{P} \}               ~~
                 [ a\neq l \land b\neq r \land hb_P=?ch(v) \land b=!ch(v) ]              
                      }
     {\ostate{10} \xrightarrow{\tau} \ostate{11}}
  $
\end{example}

The first open transitions listed, $OT_1$ and $OT_3$, only involve
possible actions from the prefix operators, respectively communication
between the two prefixes, if actions $a$ and $b$ are matching emit and
receive actions on the same channel; or simple propagation of action
$a$ from the left prefix, with no special constraint other than being
different from the local $l$ action of pLTS $C_1$.

Remark the structured states in the conclusion of these rules: both
start in state $\ostate{00}$, where both $C_1$ and $C_2$ are in their
initial states. They end in different structured states, because in
$OT_1$ both pLTS chage states, meaning that the control has gone
respectively to processes $P$ and $Q$, consuming actions $A$ and $B$,
while in $OT_3$ only $C_1$ has moved.

The last open transition $OT_{11}$ starts in state $\ostate{10}$, with
control at $P$ on the left, and action $b$ on the right. So these two
can communicate, e.g. here if $b$ is an emission, and the action of hole
$P$ is the matching reception.

@@

%
\begin{figure}[tb]
\begin{mathpar}
  \small
  \inferrule
    {\inferrule
      {
          \inferrule
              {1 \xrightarrow {l}_{C_1} 1}
              {C_1
                \models
                \openrule
                      {1 \xrightarrow{l}_{C_1} 1}
                      {\ostate{1}\xrightarrow{1}\ostate{1}}
              }
        }
        {\textrm{a:P}
          \models
            \openrule{
              1 \xrightarrow {l}_{C_1} 1,\,
              \{\xrightarrow{b_P}_P\},\, b_P\neq l}
                      {\ostate{1}\xrightarrow{b_P}\ostate{1}}
        }\\
%      \sm{fresh}v         \\
      \inferrule%*[right={L_1}]
        {
          \inferrule
              {0 \xrightarrow{b}_{C_2} 1}
              {C_4
                \models
                \openrule
                      {0 \xrightarrow{b}_{C_2} 0,\, b\neq r}
                      {\ostate{0}\xrightarrow{b}\ostate{1}}
              }
        }
        {
          \textrm{b:Q}\models
              \openrule
                  { 0 \xrightarrow {b}_{C_2} 1,\, \ b\neq r}
                  {\ostate{0}\xrightarrow{b}\ostate{1}}
        }   
    }
    {
     \textrm{PN1}
     \models
        \openrule{
                       1 \xrightarrow{l}_{C_1} 1 ~~
                       0 \xrightarrow{b}_{C_2} 1 ~~
                       \{ \xrightarrow{b_P}_{P} \}               ~~
                 [ a\neq l \land b\neq r \land b_P=?ch(v) \land b=!ch(v) ]              
                      }
     {\ostate{10} \xrightarrow{\tau} \ostate{11}}
      }\vspace{-4ex}
\end{mathpar}
  \caption{Proof of  $OT_{11}$ (with interaction of processes $P$ and action $b$) for ``a.P||b.Q''}
  \label{usingrules:OT11}
\end{figure}

\TODO{Change the example...?}
\begin{example} \emph{Using the operational rules to compute
    open-transitions}
  In Fig. \ref{usingrules:OT11} we show the deduction tree used to construct and prove the 
  open transition $OT_{11}$ of \texttt{CCS} (see example page
  \pageref{OT:ccs-composed}).
%  \TODO{Eric: change for a deduction tree of CCS expression}
The proof tree uses TR1 twice, for the $l$ transition of $C_1$ and for the
  $b$ transition of $C_2$, then uses an action $b_P$ of hole $P$, and
  combines the results using the second vector of the PN2 sub-pNet, and
  the third vector of the top node. This yields a final $\tau$
  transition. The deduction tree in the figure shows how the
  predicates are generated in this process.

@@

@@

\subsection{Presentation of algebra}
One of the important goal of the Open pNet framework is to provide us
with a model able to express the semantics of many different algebras
or languages. In particular, as we have seen in section
\ref{section:pnets}, the \emph{term algebra} used to express the
action expressions and the data in their parameters is left open in
the pNet formalisation.

@@

An algebra presentation contains:
\begin{itemize}
	\item Sorts: Constants sets of the algebra, types of the data
          and actions. 
	\item Operators: Operators used to construct action
          expressions (including their data parameters); they are
          constructors of data and action sorts, but also predicates
          over these objects.
        \item Constants: The satisfiability solver need to know which
          actions in the pLTS behaviors are constants (like the $l$,
          $r$ local actions of the controllers in the CCS example). So
          for a given pNet system, the Action sort will include all
          these constants.
\end{itemize}

\paragraph{Sorts}
Ownership of the variables and constants in a pNet acting as actions
or data is probably various. Sorts are represented by their name. The
sort of actions is built by the constant actions. Sorts here act as
sets of actions for pNet, pLTS or 
holes. There \emph{must} be one sort for actions, and eventually
several others for the types of data parameters. Sorts for integer
and boolean are implicitly declared.

\QIN{
In the CCS example, we have a sort for actions called $Action$:
}

\[ Action = \{ l, r\} \]

\paragraph{Operators}
We need introduce operators for building expressions, of both action
and data sorts, when encoding the 
algebra into pNets. So that is needed to know about the input and
output of the operators used in the expressions from those
algebra. The input and output should belong to the sorts of the
pNet. When there are $k$ sorts in this pNet, the structure of a
$n$-ary operator is vector as $< Sort_{i_1}, ... , Sort_{i_n} >
\rightarrow Sort_j , i_1 , ... , i_n, j \in [0..k] $.

\QIN{For instance, in our CCS running example, there is the action expression $!ch(v)$ using an operator $!$ presenting the ``$EMIT$'' action in CCS. We only use a sort called $Action$ to involve all the constant actions or action expressions in pNets and other parameters such as channel $ch$ or variable $v$ are all integers. So here we treat $EMIT$ as a binary operator with the format like:
}

 \[ < Int, Int > \rightarrow Action . \]

 \paragraph{Implementation of the expression meta-model}
 \TODO{maybe we could remove this paragraph (and the figure)}
The VerCors platform uses Ecore and EMF to define the structure of
both pNets and actions algebras. There exists a predefined set of
action and data expressions classes, including boolean, integer and
action expressions, with predefined variables, literal (constants),
and operators. Operators can be categorized into unary, binary and
N-ary operators.
The original version of VerCors did not allow to extend these classes,
we have modified its architecture so that a user can extend them,
defining more elements for his specific algebra.

In Fig~\ref{schema:expression-pnets}, we show the Expression class we
improved, including extensions for CCS and LOTOS.

@@

\begin{figure}[t]
  \includegraphics[width=\linewidth]{XFIG/Expression}
  \caption{The architecture of Expression class}  \label{schema:expression-pnets}
\end{figure}

@@

In our prototype we start directly with open pNets expressed
programmaticaly, using the pNet API (graphical editors may be added
later). From these, we compute
the open automaton of an open pNet by directly
implementing the sketch showed before. However, we made some
modifications due to implementation choices:
\begin{itemize}
  \item Steps 2a-2b are merged, applying Tr2 from the
premisses OTs from subnets to generate OTs without
explicitely constructing the deduction tree.
  \item We define a general naming schema for implementing the fresh 
and clone functions with more structure than a simple global name
generator. It ensures unicity, but builds names
readable enough for debugging purpose.
  \item Besides that, we also have designed a management of state variables and
    assignments throughout the whole computation.
%%   \item Satisfiability check is done only at
%% top level of the open-transition construction. An alternative would be
%% to submit the satisfiability check to the SMT solver at each level of
%% the hierarchical OT construction, potentially reducing the overall number of
%% combinations. But the submission to the SMT engine is costly, and more
%% complexity analysis is required before deciding if this would be
%% worthwhile. 
\end{itemize}

In the following sections, we detail:
\begin{itemize}
\item The structure of fresh variable names.
\item The details of the implementation for step 2a-2b, refined into:
  \subitem combining all possibilities of subnets activity/inactivity,
  \subitem matching each combination with synchronization vectors.
\item Management of local variables of pLTSs.
\item Pruning unsatisfiable OTs using the Z3 SMT solver.
\end{itemize}

@@

The variables in each synchronization vector are considered local, so
does the variables of an open transition such as its hole behaviors
and its result action. So we want rename the variables to make them
unique on every occurrence of a vector. Similarly, the name given to
the hole behaviors for each hole or the result action for each open
transition must be distinct. At the same time, we also hope them still
readable. Here we have defined a structure for the name of the fresh
variables, the variables renamed with a regular format by the function
fresh(). 
\TODO{Could try to reduce the text, but give a couple examples ?}

@@

\begin{definition}[Fresh variable]\label{fresh-variable}
The format of \emph{fresh variable} (renaming) is defined as:
$$\emph{prefix : tree\ index\ :\ counter}$$
\begin{itemize}
   \item[$\bullet$] $prefix$ is a default internal name for the variable. So far the internal name can be  sva (SV action), ra (result action), hb (hole behavior).
   \item[$\bullet$] $tree\ index$ is the index of the node in the tree-like structure of the pNet.
   \item[$\bullet$] $counter$ is the current value of the corresponding counter.
\end{itemize}
\end{definition}

@@

$Prefix$ avoids the confusion between variables from different structures with the same name by attaching the type of the structure. $Tree\ index$ is given to every node of the pNets to mention which node the variable belongs to as pNets has a tree-like structure. To each node, the tree index is always a number sequence of the tree index of its higher level and the index of the node in this level at the last. $A\ set\ of\ counters$ is used to count the current times the SV, Hole, subnet invoked, or the number of possible OT generated to provide an identity.

@@

The predicate of the open transition is a conjunction of two parts
every time applying Tr2. One part composes the predicate from the
subnets and the guard from transitions in sub-pLTS or synchronization
vectors. The other part is the conjunction of equations generated
during matching the subnets' behaviors with the synchronization
vectors. Note that it is simple to collect required terms in the first
part while in the second part we need to find out all the possible
conjunction forms. We present here a algorithm divided in two
step. Combining enumerates all the possible combinations of the
working status. Matching synchronizes the subnets according to
synchronization vectors to generate possible predicates.

%Combining is to generate all the possible matchings of the subnets' OTs (open transitions). The result list of the matched tuples called *combination*.
% 1. Initialize the result list of OT tuples.
% 2. Get one list of OTs. It's same as get OTs of one subnet.
% 3. Add *null* into the list, it means this subnet does not work this time.
% 4. Combine the rest subnets' OTs. Get the combination of the rest OTs.
% 5. If there are OTs of the chosen list not be chosen, choose one OT.
%      1. If there are tuples of the combination not be chosen, choose one tuple.
%          1. Add the chosen OT into the tuple.
%          2. Add the tuple into the result list, then do the recursion.
% 6. Return the result list.
\paragraph{Combining}

We first conduct combining. For each subnet, we enumerate all the
possible cases of its working status. The result action of a pLTS or a
pNet is the external action of the node shows the working status of
it.

>>>>

\QIN{
We choose the LOTOS formula P>>(acc(v);Q), composing several operators of LOTOS language.
}

\begin{figure}[h]
  \centerline{\includegraphics[width=8cm]{XFIG/LOTOSFormula.eps}}
  \caption{pNet encodings for the LOTOS formula}  \label{schema:lotos-pnet}
\end{figure}

@@

\QIN{
In Fig.~\ref{schema:lotos-pnet}, we show the encoding of the formula using the encoding.
}

\TODO{To be re-written[DONE]}

\QIN{
There are two styles of encodings used in the formula encoding. The first style is ``data oriented'', meaning that the state of the node controller, here is the pLTS C1, is encoded in the value of the pLTS state variable(s). Here we have a single state variable s0, initialized as 0. Its value is ``0'' when the control is on P, and ``1'' when the control is on Q. The second one uses a controller with two states to indicate the position of the control point in the right side of the formula $(acc(v);Q)$. So it is ``state oriented''. That is how we show prefix operator in PN2. 
}
\QIN{
The transitions in the pLTS have the more general syntax: ``<action-expr> [guard] {sequence of assignments}''. Guards and assignments can only use the values of variables of their source state and values from input variables in the action expression; and a transition can only assign variables of its target state. 
}

\QIN{
Each side of the formula is separately act as a hole $P$ and a sub-pNet PN2. Three synchronization vectors are given for synchronizing the behavior of the subnets to make the pNet act as the Enable operator.. The first vector, transmitting an action $a_1$ of P, can only occur if a1 is not a $\delta$, the vector concrete syntax is typically ``$< l, a_1 , \_ > -> a_1$'', in which ``$\_$'' means that the corresponding subnet is not involved in this synchronization. The second one composes the action from the subnets and gives out a synchronized action $\underline{\delta(x)}$, where the underlined action stands for an internal (already synchronized) action, that cannot be further synchronized at upper levels. Synchronized action is a straightforward generalization of the notion of internal actions, that will be convenient for observing internal events during model-checking. The second vector constraints that the transition occurs only when process $P$ performs a $\delta(x)$ action, transferring the control to $Q$, and sending a value $x$ at the same time. This value must be ``accepted''  through an $acc(x)$ action before $Q$ taking the control. The last one acts similar with the the first one, it transmits the action $a_2$ of Q.
}

\QIN{
The sub-pNet PN2 also have two synchronization vectors to make it act as the behavior of Prefix operator.
}

\QIN{
It should be easy to see that this pNet system encodes properly the behavioral semantics of the corresponding CCS expression. Remark that it is built in a systematic structural way, one pNet node encoding each CCS operator. And it is parameterized both at the level of actions variables (occurring within the controllers and the synchronization vectors), and at the level of process parameters (holes).
}

@@

%The preceding definition relies on the auxiliary functions below:

\TODO{we do not use the Sort in the implementation of synch vector
  matching. May be we could remove it from here, but it will not gain
  much...[DONE]}

%\begin{definition}[Sorts, Holes, Leaves of pNets]
%  \begin{itemize}
%  \item The sort of a pNet is its signature, i.e. the set of actions it can
%perform. In the definition of sorts, we do not need to distinguish
%input variables (that specify the dataflow within LTSs), so for
%computing LTS sorts, we use a substitution operator\footnote{$\subst{y_k\gets x_k}^{k\in K}$ is the parallel substitution 
%operation.} to remove the
%\emph{input marker} of variables. Formally:
%\[
%\begin{array}{l}
%\Sort(\mylangle S,s_0, \to\myrangle) = \{\alpha\subst{x \gets ?x| 
%x\in\symb{iv}(\alpha)}|s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
%		J}\rangle} s'\in \to \} \\ \Sort(\mylangle
%\set{\pNet}, \set{\symb{S}}, \set{\symb{SV}}\myrangle) =\{\alpha'_k |\, \alpha_j^{j\in J_k}\to\alpha'_k\in\set{\symb{SV}}\}
%\end{array}
%\]
%
%\item
%The set of holes of a pNet is defined inductively; the sets of holes
%in a pNet node and its subnets are all disjoint:
%  \[\begin{array}{l}
%\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset \\
%\Holes(\mylangle \pNet_i^{i\in I}\!,S_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) =J\cup{\displaystyle \bigcup_{i\in I}\Holes(\pNet_i)}\\
%\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
%\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  \Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
%\end{array}\]
%\item
%The set of leaves of a pNet is the set of all pLTSs occurring in the structure, defined inductively as:
%\[\begin{array}{l}
%\Leaves(\mylangle S,s_0, \to\myrangle) \!=\! \{ \mylangle S,s_0, \to\myrangle \}\\
%\Leaves(\mylangle \pNet_i^{i\in I}\!,S_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) = {\displaystyle \bigcup_{i\in I}\Leaves(\pNet_i)}
%\end{array}\]
%\end{itemize}
%\end{definition}
%
%A pNet $Q$ is \emph{closed} if it has no hole: $\Holes(Q)=\emptyset$; else it
%is said to be \emph{open}.

@@

\QIN{[ change the example OT here ]}
\begin{example}\emph{An open-transition.}
  \label{OT:lotos-composed}
\TODO{: change for an OT of new Lotos example, may be only one or two
  of them ?[DONE]}
The \texttt{LOTOS} pNet of Fig. \ref{schema:lotos-pnet} has 1 controller and 1 hole. We show its full open Automaton in section \ref{section:full-result}, it has 5 open transitions. We detail 1 of them here:

 \smallskip
 $ ot_2  = \openrule{
                       \{0 \xrightarrow{\delta}_{C_1} 0, 0 \xrightarrow{acc(x)}_{C_2} 1\}, ~~
                       \{\xrightarrow{hb_{12}}_P\}, ~~ 
                        [s_0=0 \!\land\! hb_{12}=\delta(x)], ~~ 
                       \{s_0 \xleftarrow{} 1\}     
                      }
     {\ostate{00} \xrightarrow{\underline{\delta(x)}} \ostate{01}}$
\end{example}

%The first open transitions listed, $OT_1$ and $OT_3$, only involve
%possible actions from the prefix operators, respectively communication
%between the two prefixes, if actions $a$ and $b$ are matching emit and
%receive actions on the same channel; or simple propagation of action
%$a$ from the left prefix, with no special constraint other than being
%different from the local $l$ action of pLTS $C_1$.
%
%Remark the structured states in the conclusion of these rules: both
%start in state $\ostate{00}$, where both $C_1$ and $C_2$ are in their
%initial states. They end in different structured states, because in
%$OT_1$ both pLTS chage states, meaning that the control has gone
%respectively to processes $P$ and $Q$, consuming actions $A$ and $B$,
%while in $OT_3$ only $C_1$ has moved.
%
%The last open transition $OT_{11}$ starts in state $\ostate{10}$, with
%control at $P$ on the left, and action $b$ on the right. So these two
%can communicate, e.g. here if $b$ is an emission, and the action of hole
%$P$ is the matching reception.

\QIN{
The open transition $ot_2$ involve the possible actions from the process $P$ and prefix operator $(acc(x);Q$. The hole behavior $hb_{12}$ must be the exit action $\delta(x)$, if $(acc(x);Q$ can capture the same value $x$.
}

\QIN{
Remark the structured states in the conclusion of these rules: both of them start in state $\ostate{00}$, where both $C_1$ and $C_2$ are in their initial states. They end in different structured states, because in $ot_2$ both pLTS conduct transitions, showing that the control has gone from processes $P$ to $(acc(x);Q$, consuming actions $\delta(x)$ and $acc(x)$.
}

@@

%
%\begin{figure}[tb]
%\begin{mathpar}
%  \small
%  \inferrule
%    {\inferrule
%      {
%          \inferrule
%              {1 \xrightarrow {l}_{C_1} 1}
%              {C_1
%                \models
%                \openrule
%                      {1 \xrightarrow{l}_{C_1} 1}
%                      {\ostate{1}\xrightarrow{1}\ostate{1}}
%              }
%        }
%        {\textrm{P}
%          \models
%            \openrule{
%              1 \xrightarrow {l}_{C_1} 1,\,
%              \{\xrightarrow{b_P}_P\},\, b_P\neq l}
%                      {\ostate{1}\xrightarrow{b_P}\ostate{1}}
%        }\\
%%      \sm{fresh}v         \\
%      \inferrule%*[right={L_1}]
%        {
%          \inferrule
%              {0 \xrightarrow{b}_{C_2} 1}
%              {C_4
%                \models
%                \openrule
%                      {0 \xrightarrow{b}_{C_2} 0,\, b\neq r}
%                      {\ostate{0}\xrightarrow{b}\ostate{1}}
%              }
%        }
%        {
%          \textrm{b:Q}\models
%              \openrule
%                  { 0 \xrightarrow {b}_{C_2} 1,\, \ b\neq r}
%                  {\ostate{0}\xrightarrow{b}\ostate{1}}
%        }   
%    }
%    {
%     \textrm{PN1}
%     \models
%        \openrule{
%                       1 \xrightarrow{l}_{C_1} 1 ~~
%                       0 \xrightarrow{b}_{C_2} 1 ~~
%                       \{ \xrightarrow{b_P}_{P} \}               ~~
%                 [ a\neq l \land b\neq r \land b_P=?ch(v) \land b=!ch(v) ]              
%                      }
%     {\ostate{10} \xrightarrow{\tau} \ostate{11}}
%      }\vspace{-4ex}
%\end{mathpar}
%  \caption{Proof of  $OT_{11}$ (with interaction of processes $P$ and action $b$) for ``a.P||b.Q''}
%  \label{usingrules:OT11}
%\end{figure}

\begin{figure}[tb]
\begin{mathpar}
  \small
  \inferrule
    {
          \inferrule
              {0 \xrightarrow {\delta}_{C_1} 0 ~~
               [ s_0=0 ]
              }
              {C_1
                \models
                \openrule
                      {0 \xrightarrow{\delta}_{C_1} 0 ~~
                       [ s_0=0 ] ~~
                       \{ s_0 \leftarrow 1\}
                      }
                      {\ostate{0}\xrightarrow{\delta}\ostate{0}}
              }
	\\
      \inferrule
        {
          \inferrule
              {0 \xrightarrow{acc(x)}_{C_2} 1}
              {C_2
                \models
                \openrule
                      {0 \xrightarrow{acc(x)}_{C_2} 1}
                      {\ostate{0}\xrightarrow{acc(x)}\ostate{1}}
              }
        }
        {
          \textrm{acc(x);Q}\models
              \openrule
                  { 0 \xrightarrow {acc(x)}_{C_2} 1}
                  {\ostate{0}\xrightarrow{acc(x)}\ostate{1}}
        }   
    }
    {
     \textrm{PN1}
     \models
        \openrule{
                        0 \xrightarrow{\delta}_{C_1} 0, 0 \xrightarrow{acc(x)}_{C_2} 1 ~~
                       \{\xrightarrow{hb_{12}}_P\} ~~ 
                        [s_0=0 \!\land\! hb_{12}=\delta(x)] ~~ 
                       \{s_0 \xleftarrow{} 1\}           
                      }
     {\ostate{00} \xrightarrow{\tau} \ostate{01}}
      }\vspace{-4ex}
\end{mathpar}
  \caption{Proof of  $ot_{2}$ (with interaction of processes $P$ and action $acc(x)$) for ``P>>(acc(x);Q)''}
  \label{usingrules:ot2}
\end{figure}

\TODO{Change the example...?[DONE]}
\begin{example} \emph{Using the operational rules to compute
    open-transitions}
  In Fig. \ref{usingrules:ot2} we show the deduction tree used to construct and prove the 
  open transition $ot_{2}$ of \texttt{LOTOS}.
%(see example page \pageref{OT:ccs-composed}).
%  \TODO{Eric: change for a deduction tree of CCS expression}
%  The proof tree uses TR1 twice, for the $l$ transition of $C_1$ and for the
%  $b$ transition of $C_2$, then uses an action $b_P$ of hole $P$, and
%  combines the results using the second vector of the PN2 sub-pNet, and
%  the third vector of the top node. This yields a final $\tau$
%  transition. The deduction tree in the figure shows how the
%  predicates are generated in this process.

@@

\QIN{
The proof tree uses TR1 twice, for the $\delta$ transition of $C_1$ and for the $acc(x)$ transition of $C_2$, then uses an action $b_{12}$ of hole $P$, and combines the results using the first vector of the PN2 sub-pNet, and the second vector of the top node according to TR2. This yields a final $\underline{\delta(x)}$ transition. The deduction tree in the figure shows how the predicates are generated in this process. 
}

@@

\subsection{Presentation of algebra}
One of the important goal of the Open pNet framework is to provide us
with a model able to express the semantics of many different algebras
or languages. In particular, as we have seen in section
\ref{section:pnets}, the \emph{term algebra} used to express the
action expressions and the data in their parameters is left open in
the pNet formalization.

@@

An algebra presentation contains:
\begin{itemize}
	\item Sorts: Constants sets of the algebra or types of the data
          and actions. 
	\item Operators: Operators used to construct action
          expressions (including their data parameters); they are
          constructors of data and action sorts, but also predicates
          over these objects.
        \item Constants: The satisfiability solver need to know which
          actions in the pLTS behaviors are constants (like the $l$, $\delta$, 
          $r$ of the controllers in running example). 
          For a given pNet system, a sort $Action$ will include all
          these constants.
\end{itemize}

\paragraph{Sorts}
\QIN{
Sorts are represented by a given name, showing the ownership of the variables and constants. The
sort of actions is built by the constant actions. There \emph{must} be one sort for actions, and eventually
several others for the types of data parameters. In the LOTOS example, we have the $Action$ sort:
}
\[ Action = \{ l, \delta, r, p, q, acc \} \]

In every presentation, sorts for integer and boolean are implicitly declared.

\paragraph{Operators}
%We need introduce operators for building expressions, of both action
%and data sorts, when encoding the 
%algebra into pNets. So that is needed to know about the input and
%output of the operators used in the expressions from those
%algebra. The input and output should belong to the sorts of the
%pNet. When there are $k$ sorts in this pNet, the structure of a
%$n$-ary operator is vector as $< Sort_{i_1}, ... , Sort_{i_n} >
%\rightarrow Sort_j , i_1 , ... , i_n, j \in [0..k] $.
\QIN{
We need introduce operators for building expressions, either action expressions or predicate terms, when encoding the algebra into pNets. The operators here are treated as functions, with the sorts of the pNet acting as the types of their input and output.
}

\QIN{For instance, in our LOTOS running example, we use a function ``$VOID$'' action to construct the action expression $\delta(x)$ and $acc(x)$. We can define the operator as a binary operator with the format like:
}

 \[VOID: \quad  < Action, Int > \rightarrow Action . \]

@@

%\paragraph{Implementation of the expression meta-model}
% \TODO{maybe we could remove this paragraph (and the figure)}
%The VerCors platform uses Ecore and EMF to define the structure of
%both pNets and actions algebras. There exists a predefined set of
%action and data expressions classes, including boolean, integer and
%action expressions, with predefined variables, literal (constants),
%and operators. Operators can be categorized into unary, binary and
%N-ary operators.
%The original version of VerCors did not allow to extend these classes,
%we have modified its architecture so that a user can extend them,
%defining more elements for his specific algebra.
%
%In Fig~\ref{schema:expression-pnets}, we show the Expression class we
%improved, including extensions for CCS and LOTOS.
%
%\begin{figure}[t]
%  \includegraphics[width=\linewidth]{XFIG/Expression}
%  \caption{The architecture of Expression class}  \label{schema:expression-pnets}
%\end{figure}

@@

In our prototype we start directly with open pNets expressed
programmaticaly, using the pNet API (graphical editors may be added
later). From these, we compute
the open automaton of an open pNet by directly
implementing the sketch showed before. However, we made some
modifications due to implementation choices.
%:
%\begin{itemize}
%  \item Steps 2a-2b are merged, applying Tr2 from the
%premisses OTs from subnets to generate OTs without
%explicitely constructing the deduction tree.
%  \item We define a general naming schema for implementing the fresh 
%and clone functions with more structure than a simple global name
%generator. It ensures unicity, but builds names
%readable enough for debugging purpose.
%  \item Besides that, we also have designed a management of state variables and
%    assignments throughout the whole computation.
%%%   \item Satisfiability check is done only at
%%% top level of the open-transition construction. An alternative would be
%%% to submit the satisfiability check to the SMT solver at each level of
%%% the hierarchical OT construction, potentially reducing the overall number of
%%% combinations. But the submission to the SMT engine is costly, and more
%%% complexity analysis is required before deciding if this would be
%%% worthwhile. 
%\end{itemize}
%
%In the following sections, we detail:
%\begin{itemize}
%\item The structure of fresh variable names.
%\item The details of the implementation for step 2a-2b, refined into:
%  \subitem combining all possibilities of subnets activity/inactivity,
%  \subitem matching each combination with synchronization vectors.
%\item Management of local variables of pLTSs.
%\item Pruning unsatisfiable OTs using the Z3 SMT solver.
%\end{itemize}

@@

%The variables in each synchronization vector are considered local, so
%does the variables of an open transition such as its hole behaviors
%and its result action. So we want rename the variables to make them
%unique on every occurrence of a vector. Similarly, the name given to
%the hole behaviors for each hole or the result action for each open
%transition must be distinct. At the same time, we also hope them still
%readable. Here we rename the variables with a regular format using the fresh function.

\QIN{
The variables in each synchronization vector are considered local, so
does the variables of an open transition, such as its hole behaviors
and its result action. We want all the variables unique on every occurrence of them. 
%Similarly, the name given to the hole behaviors for each hole or the result action for each open transition must be distinct. 
At the same time, we also hope them still
readable. Here we rename the variables with a regular format using the fresh function.
}
\TODO{Could try to reduce the text, but give a couple examples ?[DONE]}

@@

\begin{definition}[Fresh variable]\label{fresh-variable}
The format of \emph{fresh variable} (renaming) is defined as:
$$\emph{prefix : tree\ index\ :\ counter}$$
\begin{itemize}
   \item[$\bullet$] $prefix$ is a default internal name for the variable. So far the internal name can be  sva (SV action), ra (result action), hb (hole behavior).
   \item[$\bullet$] $tree\ index$ is the index of the node containing the variable in the tree-like structure of the pNet.
   \item[$\bullet$] $counter$ is the current value of the corresponding counter.
\end{itemize}
\end{definition}

@@

%$Prefix$ avoids the confusion between variables from different structures with the same name by attaching the type of the structure. $Tree\ index$ is given to every node of the pNets to mention which node the variable belongs to as pNets has a tree-like structure. To each node, the tree index is always a number sequence of the tree index of its higher level and the index of the node in this level at the last. $A\ set\ of\ counters$ is used to count the current times the SV, Hole, subnet invoked, or the number of possible OT generated to provide an identity.

\QIN{
For example, in the running example, the first possible behavior of the hole, it doesn't have a name, only have the prefix ``hb''. The hole is the second subnets of the root node so its ``tree index'' is ``12''. The fresh variable name is ``$:hb:12:1$''.
}

@@
